2D Yang--Mills \\ \& Cohomological Field Theory

### 2021-01-10 alpha/YangMills2D/ym2d.tex
Update YM2 note with general cohomological theory

%\addtocounter{section}{-1}
\section*{Main References}
	This review is almost entirely based on the following references:
	\begin{enumerate}[
		noitemsep
		,labelindent=\parindent
		,align=left
		,leftmargin=*
	]
	\item[\cite{Cordes:1994fc}] \textit{Cordes, Moore, Ramgoolam}, \arxiv{hep-th/9411210}
	\item[\cite{Witten:1991we}] Witten, \textit{On Quantum Gauge Theories in Two Dimensions}, 1991
	\end{enumerate}
	It is basically a condensed, more pedagocial version of \cite{Cordes:1994fc}. Many basic facts in this review are therefore uncited to avoid repeated citations of \cite{Cordes:1994fc}. All uncited claims, unless otherwise specified, can be traced back to \cite{Cordes:1994fc}. However, it is highly possible that I, in my infinite stupidity, misunderstood some ideas from \cite{Cordes:1994fc}; so please feel free to point out my mistakes.
\section{Introduction}
	We start by writing down the usual Yang--Mills action in 2D (\YM), in Euclidean signature:
	\begin{equation}
		I_{\YM}
		= +\frac{1}{4e^2}
			\int_{\Sigma_T} \dd[2]{x}
			\sqrt{G}
			\,\Tr \pqty{F_{\mu\nu} F^{\mu\nu}},
	\quad
		\sqrt{G} = \sqrt{\det G_{\mu\nu}}
	\end{equation}
	Here we will try follow the convention of \cite{Cordes:1994fc}, despite the fact that it is, unfortunately, not quite self-consistent. $\Sigma_T$ stands for the 2D \textit{target}; in the large $N$ limit, it is possible to realize \YM as a string theory with worldsheet $\Sigma_W$, as is proposed by D.~Gross and W.~Taylor, among others \cite{Gross:1992tu,Gross:1993hu,Gross:1993yt}.

@@

	The advantage of this formulation is that the $e^2 \to 0$ limit becomes non-singular; in fact, now we can simply set $e^2 = 0$, and get:
	\begin{equation}
		I[\phi,A]
		\longto I_0[\phi,A]
		= \int i \Tr\,(\phi F)
	\label{eq:top_action}
	\end{equation}
	This action is in fact \textit{topological}; there is no explicit metric dependence in the action. The measure $\mu$ comes with $e^2$ in the $I[\phi,A]$ action; setting $e^2 \to 0$ eliminates the metric dependence. Integrating out $\phi$ fixes $F = 0$, i.e.~we need only sum over the moduli of \textit{flat connections}. For a principal $G$ bundle $P\to\Sigma_T$, this is given by:
	\begin{equation}
		\mcal{M}_0
		= \mcal{M}\pqty{F = 0, P\to\Sigma_T}
		= \Bqty{
				A\in \mcal{A}(P)
				\,\Big|\,
				F(A) = 0
			} \Big/ \mcal{G}(P)
		\ \subset\ %
		\mcal{A}(P) \Big/ \mcal{G}(P)
	\end{equation}
	\vspace{-1.2\baselineskip}
	\begin{gather}
%	\begin{aligned}
		\mcal{A}(P)
%		&
		= \Bqty\Big{
			\text{all possible connections $A$ on $P$}
		} \notag
	\\
		\mcal{G}(P)
%		&
		= \Bqty\Big{
			\text{all possible gauge transformations on $P$}
		} \notag
%		\\
%	\end{aligned}\notag
	\end{gather}

@@

	Following \cite{Cordes:1994fc}, we will first review the exact solution of \YM, and then try to generalize some aspects for a generic cohomological field theory.
%	
%	In chapter 4 we relate the Hilbert space of class functions to some simple conformal field theories. The main goal is to show how bosonization leads to a natural interpretation of the Hilbert space in terms of string states.
%	
%	4.7
%	
\section{Exact Solution of 2D Yang Mills}
\subsection{Canonical Quantization on the Cylinder}
	One can perform the usual canonical quantization with $I_\YM$ on the cylinder, with coordinates $(x^0,x^1) = (t,x) \in \mbb{R}^1\times S^1$. We shall make full use of the gauge redundancies in 2D; recall that a generic gauge transformation can be written as:
	\begin{equation}
	\begin{aligned}
		A'_\mu
		&= g A_\mu g^{-1}
			+ g\,\pdd{\mu} (g^{-1}) \\
		&\simeq A_\mu
			+ D_\mu\circ\lambda,
	\end{aligned}
	\qquad
		A_\mu = A^a_\mu(t,x),
	\quad
		g = e^{-\lambda^a(t,x)\,T_a},
	\end{equation}
	\vspace{-1.2\baselineskip}
	\begin{gather}
		D_\mu = \pdd{\mu} + A_\mu^a\,T_a,
	\\[1ex]
		\pqty{D_\mu\circ\lambda}^a
		= \pdd{\mu} \lambda^a
			+ A_\mu^b\,f\id{^a_{bc}} \lambda^c,
	\quad
		(T_b)\id{^a_c} = f\id{^a_{bc}}
	\end{gather}
	Here $D_\mu$ is the $\mfrak{g}$-valued covariant derivative; it acts on the $\mfrak{g}$-valued gauge parameters $\lambda^a(t,x)$ by the adjoint representation $(T_b)\id{^a_c} = f\id{^a_{bc}}$. 
	It is thus possible to choose the \textit{temporal gauge} $A_0 = 0$, by simply solving a first order ODE of $\lambda^a(t,x)$, with respect to the variable $x^0 = t$.

@@

	By the Peter--Weyl theorem for
$G$ compact we can decompose the Hilbert space by the unitary irreducible
representations (\textit{irreps}) $\{R\}$ of $G$. 
	Consequently, a natural basis for the Hilbert space is provided by the \textit{characters} $\{\chi_R\}$ of the irreps; thus we have, relative to the $W$ basis,
	\begin{equation}
		\braket{W}{R}
		= \chi_R (W)
		= \Tr_R (W)
	\end{equation}
	This is the wave function of the $R$ basis.
	Here $\Tr_R$ denotes trace with respect to the irrep $R$; recall that $W$ is $\mfrak{g}$-valued and do not have a well-defined trace until we specify a irrep, in this case $R$. This is the gauge-invariant Wilson loop operator. Note that we are considering pure \YM without matter; the irrep $R$ is not some \textit{a priori} matter representation, but arises naturally as we consider class functions on $G$. \sidenote{This is kind of mysterious; does it has a physical interpretation, e.g.~like sum over possible matter representations?}
\subsection{Hamiltonian and the Dynamics}
	Let us now find the Hamiltonian for this theory. We are familiar with the the Hamiltonian density in 4D: $
		\frac{e^2}{2} \pqty\big{
			\vec{E}^2 + \vec{B}^2
		}
	$; here in 2D, we do not have the magnetic field $\vec{B}$, just a one component $\mfrak{g}$-valued $E = E^a T_a = F_{10}$. Therefore,
	\begin{gather}
		H = \frac{e^2}{2} \int \dd{x} \Tr E^2
		= \frac{e^2}{2} \int \dd{x}
			\delta^{ab}
			\fdv{A_1^a(x)}
			\fdv{A_1^b(x)}
		\to \frac{1}{2}\,e^2 L
			\Tr \pqty{W \pdv{W}}^2,
	\\[1ex]
		E_a = \fdv{A^a_1(x)}
		\to T_a W \pdv{W},
	\quad
		\Tr\,(T_a T_b) = \delta_{ab}
	\end{gather}
	Note that we've followed a more convenient trace convention. 
	In the last step we've restricted to the physical Hilbert space with $\Psi = \Psi(W)$. Consider $H$ action on the basis $
		\braket{W}{R}
		= \chi_R (W)
		= \Tr_R (W)
	$, we have:
	\begin{equation}
		E_a\, \chi_R (W)
		= E_a \Tr_R (W)
		= \Tr_R (T_a W),
	\end{equation}
	\vspace{-.9\baselineskip}
	\begin{equation}
	\begin{aligned}
		H\, \chi_R (W)
		&= \frac{1}{2}\,e^2 L
			\Tr_R \pqty\big{\delta^{ab} T_a T_b W} \\
		&= \frac{1}{2}\,e^2 L\,
			\pqty\big{\delta^{ab} T_a T_b}_R
			\Tr_R (W) \\
		&= \frac{1}{2}\,e^2 L
			C_2(R)\,\chi_R (W),
	\end{aligned}
	\end{equation}
	\vspace{-.3\baselineskip}
	\begin{equation}
		C_2(R)
		= \pqty\big{\delta^{ab} T_a T_b}_R
	\end{equation}
	Here $C_2(R)$ is the
\textit{quadratic Casimir} of the irrep $R$; it is proportional to $\idty_R$, or can be treated as a $c$-number when restricted to the irrep. We find out that:
	\begin{empheq}{equation}
		\textbf{The Hamiltonian $H$ is diagonalized in the $R$ basis and $\propto C_2$}
	\end{empheq}
	\vspace{-.3\baselineskip}
	\begin{equation}
		H = \frac{1}{2}\,e^2 L C_2
	\end{equation}
\subsection{Basic Amplitudes}
	Having diagonalized the Hamiltonian, we can immediately write down the $W$ basis propagator by summing over the $R$ basis; we have:
	\begin{gather}
	\begin{aligned}
		\mel*{W_1}{e^{-HT}}{W_2}
		&= \sum_R
			\braket{W_1}{R\,}
			e^{-H(R)\,T}
			\braket{R}{W_2} \\
		&= \sum_R
			\chi_R(W_1)\,
			\chi_R(W_2^\dagger)\,
			e^{-\frac{1}{2}\,e^2 a\,C_2(R)}
		= Z(W_1,W_2;a),
	\end{aligned}
	\\[.3ex]
		a = LT,
	\quad
		\chi_R^\dagger(W)
		= \pqty{\Tr_R W}^\dagger
		= \Tr_R (W^\dagger)
		= \chi_R(W^\dagger)
	\end{gather}
	This is the \textbf{cylinder amplitude}. 
	Note that the combination
$\frac{1}{2} e^2 a$ enters
together, where $a = LT$ is the area of the cylinder; this is evident from the action $I[\phi,A]$ in \eqref{eq:convenient_action}, where $e^2 \mu$ comes together as a parameter. The area dependence is also evident from the scalar action \eqref{eq:scalar_action}, where we that the theory is invariant under \textit{area preserving diffeomorphism}, $\mop{SDiff}(\Sigma_T)$. From now on we will absorb $\frac{e^2}{2}$ into $a$, and it will be the only dimensionful parameter of the theory.

@@

>>>>

%\addtocounter{section}{-1}
\section*{Main References}
	This review is almost entirely based on the following references:
	\begin{enumerate}[
		noitemsep
		,labelindent=\parindent
		,align=left
		,leftmargin=*
	]
	\item[\cite{Cordes:1994fc}] \textit{Cordes, Moore, Ramgoolam}, \arxiv{hep-th/9411210}
	\item[\cite{Witten:1991we}] Witten, \textit{On Quantum Gauge Theories in Two Dimensions}, 1991
	\end{enumerate}
	It is basically a miniaturized, reorganized, pedagocial subset of \cite{Cordes:1994fc}, focusing on its section 3, with brief summary of its section 9, 10 and 14. Many basic facts in this review are therefore uncited to avoid repeated citations of \cite{Cordes:1994fc}. All uncited claims, unless otherwise specified, can be traced back to~\cite{Cordes:1994fc}. However, it is highly possible that I, in my infinite stupidity, misunderstood some ideas from~\cite{Cordes:1994fc}; so please feel free to point out my mistakes.
\section{Introduction}
	We start by writing down the usual Yang--Mills action in 2D (\YM), in Euclidean signature:
	\begin{equation}
		I_{\YM}
		= +\frac{1}{4e^2}
			\int_{\Sigma_T} \dd[2]{x}
			\sqrt{G}
			\,\Tr \pqty{F_{\mu\nu} F^{\mu\nu}},
	\quad
		\sqrt{G} = \sqrt{\det G_{\mu\nu}}
	\end{equation}
	Here we will try follow the convention of \cite{Cordes:1994fc}, despite the fact that it is, unfortunately, not quite self-consistent. $\Sigma_T$ stands for the 2D \textit{target}; in the large $N$ limit, it is possible to realize \YM as a string theory with worldsheet $\Sigma_W$, as is proposed by D.~Gross and W.~Taylor, among others \cite{Gross:1992tu,Gross:1993hu,Gross:1993yt}.

@@

	The advantage of this formulation is that the $e^2 \to 0$ limit becomes non-singular; in fact, now we can simply set $e^2 = 0$, and get:
	\begin{equation}
		I[\phi,A]
		\longto I_0[\phi,A]
		= \int i \Tr\,(\phi F)
	\label{eq:top_action}
	\end{equation}
	This action is in fact \textit{topological}; there is no explicit metric dependence in the action. The measure $\mu$ comes with $e^2$ in the $I[\phi,A]$ action; setting $e^2 \to 0$ eliminates the metric dependence. Integrating out $\phi$ fixes $F = 0$, i.e.~we need only sum over the moduli of \textit{flat connections}. For a principal $G$ bundle $P\to\Sigma_T$, this is given by:
	\begin{equation}
		\mcal{M}_0
		= \mcal{M}\pqty{F = 0, P\to\Sigma_T}
		= \Bqty{
				A\in \mcal{A}(P)
				\,\Big|\,
				F(A) = 0
			} \Big/ \mcal{G}(P)
		\ \subset\ %
		\mcal{A}(P) \Big/ \mcal{G}(P)
	\label{eq:flat_moduli}
	\end{equation}
	\vspace{-1.2\baselineskip}
	\begin{gather}
%	\begin{aligned}
		\mcal{A}(P)
%		&
		= \Bqty\Big{
			\text{all possible connections $A$ on $P$}
		} \notag
	\\
		\mcal{G}(P)
%		&
		= \Bqty\Big{
			\text{all possible gauge transformations on $P$}
		} \notag
%		\\
%	\end{aligned}\notag
	\end{gather}

@@

	Following \cite{Cordes:1994fc}, we will first review the exact solution of \YM, and then try to generalize some aspects for a generic cohomological field theory.
%	
%	The abstract discussion concludes in section 14.6 where we give a unified description of an arbitrary cohomological field theory.
%	
%	The principles of chapter 14 are illustrated in a standard example in chapters 15, on topological Yang-Mills.

\section{Exact solution of 2D Yang--Mills}
\subsection{Canonical quantization on the cylinder}
	One can perform the usual canonical quantization with $I_\YM$ on the cylinder, with coordinates $(x^0,x^1) = (t,x) \in \mbb{R}^1\times S^1$. We shall make full use of the gauge redundancies in 2D; recall that a generic gauge transformation can be written as:
	\begin{equation}
	\begin{aligned}
		A'_\mu
		&= g A_\mu g^{-1}
			+ g\,\pdd{\mu} (g^{-1}) \\
		&\simeq A_\mu
			+ D_\mu\circ\lambda,
	\end{aligned}
	\qquad
		A_\mu = A^a_\mu(t,x),
	\quad
		g = e^{-\lambda^a(t,x)\,T_a},
	\end{equation}
	\vspace{-1.2\baselineskip}
	\begin{gather}
		D_\mu = \pdd{\mu} + A_\mu^a\,T_a,
	\\[1ex]
		\pqty{D_\mu\circ\lambda}^a
		= \pdd{\mu} \lambda^a
			+ A_\mu^b\,f\id{^a_{bc}} \lambda^c,
	\quad
		(T_b)\id{^a_c} = f\id{^a_{bc}}
	\end{gather}
	Here $D_\mu$ is the $\mfrak{g}$-valued covariant derivative; it acts on the $\mfrak{g}$-valued gauge parameters $\lambda^a(t,x)$ by the adjoint representation $(T_b)\id{^a_c} = f\id{^a_{bc}}$. 
	It is thus possible to choose the \textit{temporal gauge} $A_0 = 0$, by simply solving a first order ODE of $\lambda^a(t,x)$, with respect to the variable $x^0 = t$.

@@

	By the Peter--Weyl theorem for $G$ compact we can decompose the Hilbert space by the unitary irreducible representations (\textit{irreps}) $\{R\}$ of $G$. 
	Consequently, a natural basis for the Hilbert space is provided by the \textit{characters} $\{\chi_R\}$ of the irreps; thus we have, relative to the $W$ basis,
	\begin{equation}
		\braket{W}{R}
		= \chi_R (W)
		= \Tr_R (W)
	\end{equation}
	This is the wave function of the $R$ basis.
	Here $\Tr_R$ denotes trace with respect to the irrep $R$; recall that $W$ is $\mfrak{g}$-valued and do not have a well-defined trace until we specify a irrep, in this case $R$. This is the gauge-invariant Wilson loop operator. Note that we are considering pure \YM without matter; the irrep $R$ is not some \textit{a priori} matter representation, but arises naturally as we consider class functions on $G$. \sidenote{This is kind of mysterious; does it has a physical interpretation, e.g.~like sum over possible matter representations?}
\subsection{Hamiltonian and time evolution}
	Let us now find the Hamiltonian for this theory. We are familiar with the the Hamiltonian density in 4D: $
		\frac{e^2}{2} \pqty\big{
			\vec{E}^2 + \vec{B}^2
		}
	$; here in 2D, we do not have the magnetic field $\vec{B}$, just a one component $\mfrak{g}$-valued $E = E^a T_a = F_{10}$. Therefore,
	\begin{gather}
		H = \frac{e^2}{2} \int \dd{x} \Tr E^2
		= \frac{e^2}{2} \int \dd{x}
			\delta^{ab}
			\fdv{A_1^a(x)}
			\fdv{A_1^b(x)}
		\to \frac{1}{2}\,e^2 L
			\Tr \pqty{W \pdv{W}}^2,
	\\[1ex]
		E_a = \fdv{A^a_1(x)}
		\to T_a W \pdv{W},
	\quad
		\Tr\,(T_a T_b) = \delta_{ab}
	\end{gather}
	Note that we've followed a more convenient trace convention. 
	In the last step we've restricted to the physical Hilbert space with $\Psi = \Psi(W)$. Consider $H$ action on the basis $
		\braket{W}{R}
		= \chi_R (W)
		= \Tr_R (W)
	$, we have:
	\begin{equation}
		E_a\, \chi_R (W)
		= E_a \Tr_R (W)
		= \Tr_R (T_a W),
	\end{equation}
	\vspace{-.9\baselineskip}
	\begin{equation}
	\begin{aligned}
		H\, \chi_R (W)
		&= \frac{1}{2}\,e^2 L
			\Tr_R \pqty\big{\delta^{ab} T_a T_b W} \\
		&= \frac{1}{2}\,e^2 L\,
			\pqty\big{\delta^{ab} T_a T_b}_R
			\Tr_R (W) \\
		&= \frac{1}{2}\,e^2 L
			C_2(R)\,\chi_R (W),
	\end{aligned}
	\end{equation}
	\vspace{-.3\baselineskip}
	\begin{equation}
		C_2(R)
		= \pqty\big{\delta^{ab} T_a T_b}_R
	\end{equation}
	Here $C_2(R)$ is the \textit{quadratic Casimir} of the irrep $R$; it is proportional to $\idty_R$, or can be treated as a $c$-number when restricted to the irrep. We find out that:
	\begin{empheq}{equation}
		\textbf{The Hamiltonian $H$ is diagonalized in the $R$ basis and $\propto C_2$}
	\end{empheq}
	\vspace{-.3\baselineskip}
	\begin{equation}
		H = \frac{1}{2}\,e^2 L C_2
	\end{equation}
\subsection{Basic amplitudes}
	Having diagonalized the Hamiltonian, we can immediately write down the $W$ basis propagator by summing over the $R$ basis; we have:
	\begin{gather}
	\begin{aligned}
		\mel*{W_1}{e^{-HT}}{W_2}
		&= \sum_R
			\braket{W_1}{R\,}
			e^{-H(R)\,T}
			\braket{R}{W_2} \\
		&= \sum_R
			\chi_R(W_1)\,
			\chi_R(W_2^\dagger)\,
			e^{-\frac{1}{2}\,e^2 a\,C_2(R)}
		= Z(W_1,W_2;a),
	\end{aligned}
	\\[.3ex]
		a = LT,
	\quad
		\chi_R^\dagger(W)
		= \pqty{\Tr_R W}^\dagger
		= \Tr_R (W^\dagger)
		= \chi_R(W^\dagger)
	\end{gather}
	This is the \textbf{cylinder amplitude}. 
	Note that the combination $\frac{1}{2} e^2 a$ enters together, where $a = LT$ is the area of the cylinder; this is evident from the action $I[\phi,A]$ in \eqref{eq:convenient_action}, where $e^2 \mu$ comes together as a parameter. The area dependence is also evident from the scalar action \eqref{eq:scalar_action}, where we that the theory is invariant under \textit{area preserving diffeomorphism}, $\mop{SDiff}(\Sigma_T)$. From now on we will absorb $\frac{e^2}{2}$ into $a$, and it will be the only dimensionful parameter of the theory.

@@

\section{From \YM to general cohomological theory}
	We've learned quite a few things from our study of \YM; for example, we see that a topological theory need not be manifestly metric independent. Also, it is possible, and sometimes more elegant, to solve the theory without a complete gauge fixing. These ideas can all be utilized for the study of a general cohomological field theory.

	From a mathematical point of view, topological field theory is the study of \textit{intersection theory} on moduli spaces using physical methods \cite{Cordes:1994fc,Witten:1990bs}. In the physical framework these moduli spaces are presented in the general form:
	\begin{equation}
		\mcal{M}
		= \Bqty{
				A\in \mcal{A}
				\,\Big|\,
				DA = 0
			} \Big/ \mcal{G}
	\end{equation}
	We've seen an example of this, from \eqref{eq:flat_moduli}.

	Note that the usual Faddeev--Popov procedure and coventional BRST cohomology requires gauge-fixing. Now we would like to develop an alternative method which does not rely on a specific gauge choice. In \YM, postponing the gauge fixing leads us to the class functions on $G$, which is more convenient to work with than the gauge-fixed $\textsl{Cartan} \,\big/ \textsl{Weyl}$. More generally, we can directly work on the space of fields $\mcal{A}$ before gauge fixing, and consider \textit{$\mcal{G}$-equivariant} quantities on $\mcal{A}$. Here $\mcal{G}$ is generally an infinite-dimensional Lie group, the group of all possible gauge transformations. This is the idea of \textit{equivariant cohomology}. 
\subsection{Differential complex from ``supersymmetry''}
	Here we will briefly illustrate the construction of a differential complex, which will be a model for equivariant cohomology. We want something that is ``BRST-like'', without going through the Faddeev--Popov gauge-fixing process. The idea is to mimic the BRST field contents by introducing \textit{ghosts}, which can be systematically constructed using differential graded algebra (DGA); the grading is the \textit{ghost number}.

	One can then think of these fields as coordinates of some \textit{superspace}, and write down actions of the \textit{superfield} \cite{Cordes:1994fc,Horne:1988yn}.  It turns out that the \textit{usual} Lie algebra cohomology of this \textit{supersymmetrized} Lie algebra $\mfrak{g}^S$ is the same as the equivariant cohomology of the original Lie algebra $\mfrak{g}$ \cite{Cordes:1994fc,Witten:1990bs}.

	Example $\mcal{N} = (2,2)$

\subsection{Localization}

	2. Fields: When we study equivariant cohomology we must choose a model for the
	cohomology. A choice of differential complex amounts to a choice of fields.
	3. Equations: Mathematically, the ultimate objects of study are intersection numbers.
	The equations are used to isolate interesting subspaces of C. One writes Poincaré duals to these subspaces using MQ representatives of Thom forms (thus introducing further fields). The equations are viewed as the vanishing of a section of a vector bundle: s = 0 where s(φ) = Dφ is the section of a vector bundle we will call the
	localization bundle: s ∈ Γ[E localization → C].

	:
